---
title: "Lab7"
author: "Alice Zheng"
date: "3/4/2020"
output: html_document
---

## ----style, echo=FALSE, message=FALSE, warning=FALSE, results="asis"-------
```{r}
library("BiocStyle")
library("knitr")
library("rmarkdown")
opts_chunk$set(message = FALSE, error = FALSE, warning = FALSE,
               cache = FALSE, fig.width = 5, fig.height = 5)
```

## ----loadairway------------------------------------------------------------
```{r}
library("airway")
```

## ----dir-------------------------------------------------------------------
```{r}
dir <- system.file("extdata", package="airway", mustWork=TRUE)
```

## ----list.files------------------------------------------------------------
```{r}
list.files(dir)
list.files(file.path(dir, "quants"))
```

## ----sampleinfo------------------------------------------------------------
```{r}
csvfile <- file.path(dir, "sample_table.csv")
coldata <- read.csv(csvfile, row.names=1, stringsAsFactors=FALSE)
coldata
```

## ----makecoldata-----------------------------------------------------------
```{r}
coldata <- coldata[1:2,]
coldata$names <- coldata$Run
coldata$files <- file.path(dir, "quants", coldata$names, "quant.sf.gz")
file.exists(coldata$files)
```

## ----tximeta, message=TRUE-------------------------------------------------
```{r}
library("tximeta")
se <- tximeta(coldata)
```

## ----lookse----------------------------------------------------------------
```{r}
dim(se)
head(rownames(se))
```

## ----summarize, message=TRUE-----------------------------------------------
```{r}
gse <- summarizeToGene(se)
```

## ----lookgse---------------------------------------------------------------
```{r}
dim(gse)
head(rownames(gse))
```

## ----sumexp, echo=FALSE----------------------------------------------------
```{r}
par(mar=c(0,0,0,0))
plot(1,1,xlim=c(0,100),ylim=c(0,100),bty="n",
     type="n",xlab="",ylab="",xaxt="n",yaxt="n")
polygon(c(45,90,90,45),c(5,5,70,70),col="pink",border=NA)
polygon(c(45,90,90,45),c(68,68,70,70),col="pink3",border=NA)
text(67.5,40,"assay(s)")
text(67.5,35,'e.g. "counts", ...')
polygon(c(10,40,40,10),c(5,5,70,70),col="skyblue",border=NA)
polygon(c(10,40,40,10),c(68,68,70,70),col="skyblue3",border=NA)
text(25,40,"rowRanges")
polygon(c(45,90,90,45),c(75,75,95,95),col="palegreen",border=NA)
polygon(c(45,47,47,45),c(75,75,95,95),col="palegreen3",border=NA)
text(67.5,85,"colData")
```

## ----loadfullgse-----------------------------------------------------------
```{r}
data(gse)
gse
```

## ----assaysgse-------------------------------------------------------------
```{r}
assayNames(gse)
head(assay(gse), 3)
colSums(assay(gse))
```

## ----rowrangesgse----------------------------------------------------------
```{r}
rowRanges(gse)
```


## ----lookseqinfo-----------------------------------------------------------
```{r}
seqinfo(rowRanges(gse))
```


## ----coldatagse------------------------------------------------------------
```{r}
colData(gse)
```

## ----gsevars---------------------------------------------------------------
```{r}
gse$donor
gse$condition
```


## ----gsevarsrename---------------------------------------------------------
```{r}
gse$cell <- gse$donor
gse$dex <- gse$condition
```

## ----renamelevels----------------------------------------------------------
```{r}
levels(gse$dex)
```

# when renaming levels, the order must be preserved!
```{r}
levels(gse$dex) <- c("untrt", "trt")
```


## ----gsedex----------------------------------------------------------------
```{r}
library("magrittr")
gse$dex %<>% relevel("untrt")
gse$dex
```



## ----explaincmpass, eval = FALSE-------------------------------------------
```{r}
gse$dex <- relevel(gse$dex, "untrt")
```

## ----countreads------------------------------------------------------------
```{r}
round( colSums(assay(gse)) / 1e6, 1 )
```

## ----loaddeseq2------------------------------------------------------------
```{r}
library("DESeq2")
```

## ----makedds---------------------------------------------------------------
```{r}
dds <- DESeqDataSet(gse, design = ~ cell + dex)
```

## --------------------------------------------------------------------------
```{r}
countdata <- round(assays(gse)[["counts"]])
head(countdata, 3)
```

## --------------------------------------------------------------------------
```{r}
coldata <- colData(gse)
```

## --------------------------------------------------------------------------
```{r}
ddsMat <- DESeqDataSetFromMatrix(countData = countdata,
                                 colData = coldata,
                                 design = ~ cell + dex)
```

## --------------------------------------------------------------------------
```{r}
nrow(dds)
keep <- rowSums(counts(dds)) > 1
dds <- dds[keep,]
nrow(dds)
```


## --------------------------------------------------------------------------
```{r}
keep <- rowSums(counts(dds) >= 10) >= 3
```


## ----meanSdCts-------------------------------------------------------------
```{r}
lambda <- 10^seq(from = -1, to = 2, length = 1000)
cts <- matrix(rpois(1000*100, lambda), ncol = 100)
library("vsn")
meanSdPlot(cts, ranks = FALSE)
```


## ----meanSdLogCts----------------------------------------------------------
```{r}
log.cts.one <- log2(cts + 1)
meanSdPlot(log.cts.one, ranks = FALSE)
```

## ----vst-------------------------------------------------------------------
```{r}
vsd <- vst(dds, blind = FALSE)
head(assay(vsd), 3)
colData(vsd)
```

## ----rlog------------------------------------------------------------------
```{r}
rld <- rlog(dds, blind = FALSE)
head(assay(rld), 3)
```


## ----transformplot, fig.width = 6, fig.height = 2.5------------------------
```{r}
library("dplyr")
library("ggplot2")

dds <- estimateSizeFactors(dds)

df <- bind_rows(
  as_data_frame(log2(counts(dds, normalized=TRUE)[, 1:2]+1)) %>%
         mutate(transformation = "log2(x + 1)"),
  as_data_frame(assay(vsd)[, 1:2]) %>% mutate(transformation = "vst"),
  as_data_frame(assay(rld)[, 1:2]) %>% mutate(transformation = "rlog"))
  
colnames(df)[1:2] <- c("x", "y")  

ggplot(df, aes(x = x, y = y)) + geom_hex(bins = 80) +
  coord_fixed() + facet_grid( . ~ transformation) 
```

 

## --------------------------------------------------------------------------
```{r}
sampleDists <- dist(t(assay(vsd)))
sampleDists
```



## --------------------------------------------------------------------------
```{r}
library("pheatmap")
library("RColorBrewer")
```

## ----distheatmap, fig.width = 6.1, fig.height = 4.5------------------------
```{r}
sampleDistMatrix <- as.matrix( sampleDists )
rownames(sampleDistMatrix) <- paste( vsd$dex, vsd$cell, sep = " - " )
colnames(sampleDistMatrix) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows = sampleDists,
         clustering_distance_cols = sampleDists,
         col = colors)
```

## --------------------------------------------------------------------------
```{r}
library("PoiClaClu")
poisd <- PoissonDistance(t(counts(dds)))
```

## ----poisdistheatmap, fig.width = 6.1, fig.height = 4.5--------------------
```{r}
samplePoisDistMatrix <- as.matrix( poisd$dd )
rownames(samplePoisDistMatrix) <- paste( dds$dex, dds$cell, sep=" - " )
colnames(samplePoisDistMatrix) <- NULL
pheatmap(samplePoisDistMatrix,
         clustering_distance_rows = poisd$dd,
         clustering_distance_cols = poisd$dd,
         col = colors)
```

## ----plotpca, fig.width=6, fig.height=4.5----------------------------------
```{r}
plotPCA(vsd, intgroup = c("dex", "cell"))
```


## --------------------------------------------------------------------------
```{r}
pcaData <- plotPCA(vsd, intgroup = c( "dex", "cell"), returnData = TRUE)
pcaData
percentVar <- round(100 * attr(pcaData, "percentVar"))
```


## ----ggplotpca, fig.width=6, fig.height=4.5--------------------------------
```{r}
ggplot(pcaData, aes(x = PC1, y = PC2, color = dex, shape = cell)) +
  geom_point(size =3) +
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  coord_fixed() +
  ggtitle("PCA with VST data")
```

## --------------------------------------------------------------------------
```{r}
library("glmpca")
gpca <- glmpca(counts(dds), L=2)
gpca.dat <- gpca$factors
gpca.dat$dex <- dds$dex
gpca.dat$cell <- dds$cell
```



## ----glmpca, fig.width=6, fig.height=4.5-----------------------------------
```{r}
ggplot(gpca.dat, aes(x = dim1, y = dim2, color = dex, shape = cell)) +
  geom_point(size =3) + coord_fixed() + ggtitle("glmpca - Generalized PCA")
```

## ----mdsvst, fig.width=6, fig.height=4.5-----------------------------------
```{r}
mds <- as.data.frame(colData(vsd))  %>%
         cbind(cmdscale(sampleDistMatrix))
ggplot(mds, aes(x = `1`, y = `2`, color = dex, shape = cell)) +
  geom_point(size = 3) + coord_fixed() + ggtitle("MDS with VST data")
```

## ----mdspois, fig.width=6, fig.height=4.5----------------------------------
```{r}
mdsPois <- as.data.frame(colData(dds)) %>%
   cbind(cmdscale(samplePoisDistMatrix))
ggplot(mdsPois, aes(x = `1`, y = `2`, color = dex, shape = cell)) +
  geom_point(size = 3) + coord_fixed() + ggtitle("MDS with PoissonDistances")
```

