---
title: "Lab 10"
author: "Alice Zheng"
date: "4/1/2020"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(maps)
library(mapdata)
library(lubridate)
library(viridis)
library(wesanderson)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
daily_report <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-31-2020.csv")) %>% 
  rename(Long = "Long_") 
    
ggplot(daily_report, aes(x = Long, y = Lat, size = Confirmed/1000)) +
    borders("world", colour = NA, fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='purple', fill='blue', alpha = 0.5) +
    labs(title = 'World COVID-19 Confirmed cases',x = '', y = '',
        size="Cases (x1000))") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```

### This map shows the confirmed COVID-19 cases in the world as of March 31st, 2020. The density of cases are depicted using dots for different countries (and different regions for larger, more heavily affected countries. As the dots increase in size, the amount of cases in the area increase as well.  



```{r, echo=FALSE, warning=FALSE, message=FALSE}
daily_report <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-27-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  filter (!Province_State %in% c("Alaska","Hawaii", "American Samoa",
                  "Puerto Rico","Northern Mariana Islands", 
                  "Virgin Islands", "Recovered", "Guam", "Grand Princess",
                  "District of Columbia", "Diamond Princess")) %>% 
  filter(Lat > 0)

ggplot(daily_report, aes(x = Long, y = Lat, size = Confirmed/1000)) +
    borders("state", colour = "black", fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='blue', fill='blue', alpha = 0.5) +
    labs(title = 'COVID-19 Confirmed Cases in the US', x = '', y = '',
        size="Cases (x1000))") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```

### This map shows the amount of confirmed COVID-19 cases in the United states as of March 31st, 2020. Dots represent the density of cases in provinces. 



```{r, echo=FALSE, warning=FALSE, message=FALSE}
mybreaks <- c(1, 100, 1000, 10000, 10000)

ggplot(daily_report, aes(x = Long, y = Lat, size = Confirmed)) +
    borders("state", colour = "white", fill = "grey90") +
    geom_point(aes(x=Long, y=Lat, size=Confirmed, color=Confirmed),stroke=F, alpha=0.7) +
    scale_size_continuous(name="Cases", trans="log", range=c(1,7), 
                        breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+")) +
    scale_color_viridis_c(option="viridis",name="Cases",
                        trans="log", breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+"))  +
# Cleaning up the graph
  
  theme_void() + 
    guides( colour = guide_legend()) +
    labs(title = "COVID-19 Confirmed Cases in the US'") +
    theme(
      legend.position = "bottom",
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#ffffff", color = NA), 
      panel.background = element_rect(fill = "#ffffff", color = NA), 
      legend.background = element_rect(fill = "#ffffff", color = NA)
    ) +
    coord_fixed(ratio=1.5)
```

### This map, similar to the one above, shows the amount of confirmed cases in the United States as of March 31st, 2020. It uses a color and size scale of points to demonstrate the density of confirmed cases across the country.


```{r, echo=FALSE, warning=FALSE, message=FALSE}
daily_report <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-27-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  group_by(Province_State) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Province_State = tolower(Province_State))

# load the US map data
us <- map_data("state")
# We need to join the us map data with our daily report to make one data frame/tibble
state_join <- left_join(us, daily_report, by = c("region" = "Province_State"))

# plot state map
ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "white") +
  geom_polygon(color = "black", fill = NA) +
  scale_fill_gradient(trans = "log10") +
  labs(title = "COVID-19 Confirmed Cases in the US'")
```

### The map shows the amount of confirmed COVID-19 cases in the United states as of March 31st, 2020, using color code to show illustrate the density of cases for each state. 


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# plot state map
ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "white") +
  geom_polygon(color = "black", fill = NA) +
  scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous"),
                         trans = "log10") +
  labs(title = "COVID-19 Confirmed Cases in the US'")
```

### The map shows the amount of confirmed COVID-19 cases in the United states as of March 31st, 2020, using a different color code to show illustrate the density of cases for each state. 


```{r, echo=FALSE, warning=FALSE, message=FALSE}
daily_report <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-27-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "Massachusetts") %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))

us <- map_data("state")
ma_us <- subset(us, region == "massachusetts")
counties <- map_data("county")
ma_county <- subset(counties, region == "massachusetts")

state_join <- left_join(ma_county, daily_report, by = c("subregion" = "Admin2")) 

# plot state map
ggplot(data = ma_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "white") +
    scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous"),
                         trans = "log10") +
  labs(title = "COVID-19 Confirmed Cases in Massachusetts'")
```

### This map shows the density of confirmed COVID-19 cases in the state of Massachusetts as of March 31st, 2020. The cases are color coded and separated into provinces. 



```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(plotly)

ggplotly(
  ggplot(data = ma_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "black") +
    scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous")) +
  ggtitle("COVID-19 Cases in MA") +
# Cleaning up the graph
  labs(x=NULL, y=NULL) +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank())
)
```

### This is an interactive map which shows the density of confirmed COVID-19 cases in the state of Massachusetts as of March 31st, 2020. The cases are color coded and separated into provinces. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
time_series_confirmed_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")  %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Confirmed") 

# Let's get the times series data for deaths

time_series_deaths_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")  %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Deaths")

time_series_recovered_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region") %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Recovered")

# Create Keys 

time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)

time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)

time_series_recovered_long <- time_series_recovered_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Recovered)

# Join tables
time_series_long_joined <- full_join(time_series_confirmed_long,
              time_series_deaths_long, by = c("Key"))

time_series_long_joined <- full_join(time_series_long_joined,
              time_series_recovered_long, by = c("Key")) %>% 
    select(-Key)

# Reformat the data
time_series_long_joined$Date <- mdy(time_series_long_joined$Date)

# Create Report table with counts
time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date),
               names_to = "Report_Type", values_to = "Counts")
```

```{r, echo=FALSE, warning=FALSE}
library(ggplot2)
library(gganimate)
library(transformr)
theme_set(theme_bw())
```



```{r, echo=FALSE, warning=FALSE, message=FALSE}
data_time <- time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths", "Recovered"), sum) %>% 
    filter (Country_Region %in% c("China","Italy","Germany","US")) 

p <- ggplot(data_time, aes(x = Date,  y = Confirmed, color = Country_Region)) + 
      geom_point() +
      geom_line() +
      ggtitle("Confirmed COVID-19 Cases") +
      geom_point(aes(group = seq_along(Date))) +
      transition_reveal(Date) 
    
animate(p,renderer = gifski_renderer(), end_pause = 15)
```

### This animation shows the increasing of confirmed COVId-19 cases in China, Italy, Germany, and the United states graphed together over time. 



```{r, echo=FALSE, warning=FALSE, message=FALSE}
covid <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
           rename(Province_State= "Province/State", Country_Region = "Country/Region") %>%
           pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                  names_to = "Date", values_to = "Confirmed") %>%
           mutate(Date = mdy(Date) - days(1),
                  Place = paste(Lat,Long,sep="_")) %>%
# Summarizes state and province information
             group_by(Place,Date) %>%
           summarise(cumulative_cases = ifelse(sum(Confirmed)>0,
                     sum(Confirmed),NA_real_),
                     Lat = mean(Lat),
                     Long = mean(Long)) %>%
           mutate(Pandemic_day = as.numeric(Date - min(Date)))

world <- ggplot(covid,aes(x = Long, y = Lat, size = cumulative_cases/1000)) +
                 borders("world", colour = "gray50", fill = "grey90") +
                 theme_bw() +
                 geom_point(color='purple', alpha = .5) +
                 labs(title = 'Pandemic Day: {frame}',x = '', y = '',
                      size="Cases (x1000))") +
                 theme(legend.position = "right") +
                 coord_fixed(ratio=1.3)+
                 transition_time(Date) +
                 enter_fade()

animate(world, end_pause = 30)
```

### This animation shows a map of the world and the increasing of confirmed cases based on country. It increases the number of points on the map as days of the pandemic increases.


# REFERENCES

Application written in R using data from [@JohnsHopkins2020].


